\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-28 paper/main.tex
update fig4 and corresponding numbers

Compared to experimental data, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher in the
absence of cilia-mediated flow. Moreover, the ROI 5 $\cbar$ profiles
for the two configurations in our computational model deviate,
which is not the case in the experiments. Altogether,
our findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for
solute movement to regions upstream in the cilia-mediated flow fields or to
distant ventricles, while cilia affect the downstream mean concentrations.

@@

%--------------VERIFICATION--------------%
\section{Verification}\label{sec:appB_verification}
\subsection{Method of Manufactured Solutions}\label{subsec:mms}
Let $\hat{\xx} = (\hat{x}, \hat{y}, \hat{z}) = \frac{1}{L}(x, y, z)$
be the spatial coordinates scaled by the length $L$, which is the length of
the computational domain $\Omega$ in the $x$-direction. Letting $\bm{\phi}$ be
the vector-valued function
\begin{equation*}
    \bm{\phi} = \left(\sin{(\pi(\hat{x} - \hat{y}))},
    \sin{(\pi(\hat{y} + \hat{z}))},
    \sin{(\pi(\hat{x} - \hat{z}))}\right)
\end{equation*}
and defining the velocity as 
\begin{equation}
    \uu_{\mathrm{mms}} = \nabla\times\bm{\phi},
    \label{eq:velocity_mms}
\end{equation}
we have by construction that the velocity is divergence-free,
owing to the identity $\nabla\cdot(\nabla\times\bm{F}) = 0$ for
an arbitrary vector field $\bm{F}$. The pressure was defined as
\begin{equation}
    p_{\mathrm{mms}} = \cos{(\pi(\hat{x} + \hat{y} + \hat{z}))}.
    \label{eq:pressure_mms}
\end{equation}
Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions},
we construct a problem with the Stokes equations where~\Cref{eq:velocity_mms}
and~\Cref{eq:pressure_mms} are the solutions.
To this end, we consider the general form of the steady Stokes equations
\begin{subequations}
    \begin{align}
        -\nabla\cdot\bsig &= \ff \qquad\mathrm{in \ }\Omega,\\
        \nabla\cdot\uu &= g \qquad\mathrm{in \ }\Omega,
    \end{align}
    \label{eq:stokes_mms}%
\end{subequations}%
with $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and
$\beps(\uu) = \frac{1}{2}\left(\nabla\uu + (\nabla\uu)^T\right)$, and the boundary condition
\begin{alignat}{2}
    \uu &= \uu_{\mathrm{mms}} \qquad&&\mathrm{on \ }\Gamma,
\end{alignat}
where $\Gamma=\partial\Omega$ is the boundary of the domain.
Since the velocity field defined by~\Cref{eq:velocity_mms} is divergence free by construction,
$g = 0$. Setting the external force $\ff=-\nabla\cdot\bsig(\uu_{\mathrm{mms}}, p_{\mathrm{mms}})$
using~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
the solutions $\uu$ and $p$ to~\Cref{eq:stokes_mms} will be the ones
defined by~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
as long as we enforce equality in the mean pressures, i.e. $\intO{(p_h - p_{\mathrm{mms}})}=0$.
Convergence rates are then readily calculated by comparing the numerical approximations of
$\uu$ and $p$ with $\uu_{\mathrm{mms}}$ and $p_{\mathrm{mms}}$
on meshes of varying grid resolution.

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated $\mathbf{L}^2$
error norms and $\mathbf{H}^1$ error semi-norms for $\uu_h$,
and $L^2$ error norms for $p_h$~(\Cref{tab:cylinder_error_rates}).
Based on the error norm calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norms of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \small
    \centering
    \caption{Error norms of the velocity $\uu_h$ in the
    $\mathbf{L}^2$ and $\mathbf{H}^1$ norms,
    the pressure $p_h$ in the $L^2$ norm,
    and the error in the maximum velocity magnitude $\uu_{h, {\mathrm{mag}}}$
    for mesh refinement on a cylinder mesh with the only-cilia/no-cardiac flow model.
    Additionally, we report the $L^2$ norm of the divergence of the velocity.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\normlinf{u_h} -\normlinf{u_{\mathrm{ex}}}$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.4e-01}  (--) 
        & \num{2.4e-02} (---) & \num{1.4e-07} \\

@@

Particle movement varied over the different measurements. The mean of $\overline{u}_x$ was
0.26 \textmu m/s and the mean of $\overline{u}_y$ was 0.08 \textmu m/s.
The maximum mean velocities
in absolute values were 10.2 \textmu m/s and 2.4 \textmu m/s for
$\overline{u}_x$ and $\overline{u}_y$,
respectively. We used the only-cardiac/no-cilia flow model to tune the
pressure boundary condition
amplitude $A$. The choice of $A=1.5$ mPa was used since: (i) it yields a
maximum velocity magnitude
of 4.8 \textmu m/s, which is close to the maximum observed in the
particle tracking data when
rejecting the outlier at around 10 \textmu m/s; (ii) it yields a time-averaged mean velocity
of 0.89 \textmu m/s, which is close to the mean observed in the particle tracking data.
\begin{figure}[H]
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_x.png}
    \end{minipage}
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_y.png}
    \end{minipage} 
    \caption{Particle tracking data. Mean velocities in
    the $x$ (\textbf{a.}) and $y$ (\textbf{b.}) directions of the particles tracked.
    The dashed horizontal line designates a value of zero.}\label{fig:particle_tracking_data}
\end{figure}

>>>>

Compared to experimental data, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 81\% and 28\% higher in the
absence of cilia-mediated flow. Moreover, the ROI 5 $\cbar$ profiles
for the two configurations in our computational model deviate,
which is not the case in the experiments. Altogether,
our findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for
solute movement to regions upstream in the cilia-mediated flow fields or to
distant ventricles, while cilia affect the downstream mean concentrations.

@@

%--------------VERIFICATION--------------%
\section{Verification}\label{sec:appB_verification}
\subsection{Method of Manufactured Solutions}\label{subsec:mms}
Let $\hat{\xx} = (\hat{x}, \hat{y}, \hat{z}) = \frac{1}{L}(x, y, z)$
be the spatial coordinates scaled by the length $L$, which is the length of
the computational domain $\Omega$ in the $x$-direction. Letting $\bm{\phi}$ be
the vector-valued function
\begin{equation*}
    \bm{\phi} = \left(\sin{(\pi(\hat{x} - \hat{y}))},
    \sin{(\pi(\hat{y} + \hat{z}))},
    \sin{(\pi(\hat{x} - \hat{z}))}\right)
\end{equation*}
and defining the velocity as 
\begin{equation}
    \uu_{\mathrm{mms}} = \nabla\times\bm{\phi},
    \label{eq:velocity_mms}
\end{equation}
we have by construction that the velocity is divergence-free,
owing to the identity $\nabla\cdot(\nabla\times\bm{F}) = 0$ for
an arbitrary vector field $\bm{F}$. The pressure was defined as
\begin{equation}
    p_{\mathrm{mms}} = \cos{(\pi(\hat{x} + \hat{y} + \hat{z}))}.
    \label{eq:pressure_mms}
\end{equation}
Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions},
we construct a problem with the Stokes equations where~\Cref{eq:velocity_mms}
and~\Cref{eq:pressure_mms} are the solutions.
To this end, we consider the general form of the steady Stokes equations
\begin{subequations}
    \begin{align}
        -\nabla\cdot\bsig &= \bm{f} \qquad\mathrm{in \ }\Omega,\\
        \nabla\cdot\uu &= g \qquad\mathrm{in \ }\Omega,
    \end{align}
    \label{eq:stokes_mms}%
\end{subequations}%
with $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and
$\beps(\uu) = \frac{1}{2}\left(\nabla\uu + (\nabla\uu)^T\right)$, and the boundary condition
\begin{alignat}{2}
    \uu &= \uu_{\mathrm{mms}} \qquad&&\mathrm{on \ }\Gamma,
\end{alignat}
where $\Gamma=\partial\Omega$ is the boundary of the domain.
Since the velocity field defined by~\Cref{eq:velocity_mms} is divergence free by construction,
$g = 0$. Setting the external force $\bm{f}=-\nabla\cdot\bsig(\uu_{\mathrm{mms}}, p_{\mathrm{mms}})$
using~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
the solutions $\uu$ and $p$ to~\Cref{eq:stokes_mms} will be the ones
defined by~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
as long as we enforce equality in the mean pressures, i.e. $\intO{(p_h - p_{\mathrm{mms}})}=0$.
Convergence rates are then readily calculated by comparing the numerical approximations of
$\uu$ and $p$ with $\uu_{\mathrm{mms}}$ and $p_{\mathrm{mms}}$
on meshes of varying grid resolution.

@@

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \small
    \centering
    \caption{Error norms of the velocity $\uu_h$ in the
    $\mathbf{L}^2$ and $\mathbf{H}^1$ norms,
    the pressure $p_h$ in the $L^2$ norm,
    and the error in the maximum velocity magnitude $\uu_{h, {\mathrm{mag}}}$
    for mesh refinement on a cylinder mesh with the cilia-driven/no-cardiac flow model.
    Additionally, we report the $L^2$ norm of the divergence of the velocity.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\normlinf{u_h} -\normlinf{u_{\mathrm{ex}}}$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.4e-01}  (--) 
        & \num{2.4e-02} (---) & \num{1.4e-07} \\

@@

Particle movement varied over the different measurements. The mean of $\overline{u}_x$ was
0.26 \textmu m/s and the mean of $\overline{u}_y$ was 0.08 \textmu m/s.
The maximum mean velocities
in absolute values were 10.2 \textmu m/s and 2.4 \textmu m/s for
$\overline{u}_x$ and $\overline{u}_y$,
respectively. We used the cardiac-induced/no-cilia flow model to tune the
pressure boundary condition
amplitude $A$. The choice of $A=1.5$ mPa was used since: (i) it yields a
maximum velocity magnitude
of 4.8 \textmu m/s, which is close to the maximum observed in the
particle tracking data when
rejecting the outlier at around 10 \textmu m/s; (ii) it yields a time-averaged mean velocity
of 0.89 \textmu m/s, which is close to the mean observed in the particle tracking data.
\begin{figure}[H]
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_x.png}
    \end{minipage}
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_y.png}
    \end{minipage} 
    \caption{Particle tracking data. Mean velocities in
    the $x$ (\textbf{a.}) and $y$ (\textbf{b.}) directions of the particles tracked.
    The dashed horizontal line designates a value of zero.}\label{fig:particle_tracking_data}
\end{figure}
